
## First load in the required packages data
```{r}
#the main package
library(FSA)
#this package contains data compatible with this package
library(FSAdata)
#for data manipulation
library(tidyverse)
library(stringr)
```

## Then let's grab the data and process it!
## We need a dataframe that contains age and catch (both are numerics)
```{r}
#can use this function from FSAdata to find data that is compatible with the function
help.search("Catch curve", package=c("FSAdata","FSA"))

#we are going to use data of estimated catch-at-age for Gulf Menhaden from 1964-2004
data<-FSAdata::Menhaden1 %>% 
  #pivot the data to get it in the required format
  pivot_longer(cols = age0:age6, names_to = "age", values_to = "ct") %>% 
  #then use group by and summarize to get one total catch value by age class
  group_by(age) %>% summarise(catch=sum(ct))

#use string manipulations to remove the leading age_ from the values in the age column
data$age<-str_replace_all(data$age, "age", "")

#convert columns to numerics
data$age<-as.numeric(data$age)

#take the log of catch
data$logct <- log(data$catch)

#examine the structure of the dataframe
str(data)
```

## Plot the catch curve for this data to assess which ages fall on the descending limb of the catch curve.
```{r}
plot(logct~age,data=data,ylab="log(Catch)",pch=19)
```

## Non-weighted regression
## Now, you can use this information in the catchCurve function
```{r}
#we will use years 1-6
results <- FSA::catchCurve(catch~age,data=data,ages2use=1:6)

#show the estimated values of Z and A (and std. deviation)
#Z is the instantaneous mortality rate (think slope of line)
#A is the annual mortality rate
summary(results)

#show the confidence intervals associated with these values
confint(results)

#plot the points with the results for Z and A
plot(results)

```

## Weighted Regression
## The use of catchCurve in the previous chunk assigned equal weight to fish of all ages, but Maceina and Bettoli (1998) suggested that a weighted regression should be used with the catch-curve method in order to reduce the relative impact of older ages with fewer fish. The use of use.weights=TRUE satisfies this.
```{r}

weighted_results <- FSA::catchCurve(catch~age,data=data,
                                    ages2use=2:6,use.weights=TRUE)

summary(weighted_results)

confint(weighted_results)

plot(weighted_results)
```

## Portfolio Question
How do the estimates of Z and A change between between the weighted and non-weighted regressions? How does the confidence interval change?





