---
title: "Portfolio 6"
author: "Kat Aristi"
date: "2024-02-04"
output: html_document
---

# Welcome to the FSA function depletion()!!!!
## This function can be used to estimate the initial population size (N0) for a closed population (aka no immigration, emmigration, mortatlity, or recruitment).
## There are three methods for estimating N0: the Leslie Method, DeLury Method, and K-Pass Method
## Before we get into that, lets load in the required packages
```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#the main package
install.packages("FSA")
library(FSA)

#the data compatible with this package
install.packages("FSAdata")
library(FSAdata)

#other necessary packages
library(tidyverse)
library(janitor)
```

## Now load in the depletion data!
### If you go to the help window and click on the depletion package, you will also find details about the Leslie and DeLury methods.
```{r}
help.search("depletion", package=c("FSAdata","FSA"))

#we are going to use the catch and effort snapper data 
#there are 3 different species within this data set, all with the same effort. We are going to focus on Pristipomoides zonatus species. 
#effort = fishing effort (line-hours of a bottom hand-line)

snapper <- FSAdata::Pathfinder
snapper_pzonatus <-snapper %>% clean_names() 
```

## Now to start with the Leslie Method!
```{r}
#grab the data use the depletion() to calculate N0 and q
leslie <- with(snapper_pzonatus, depletion(pzonatus,effort,method="Leslie",Ricker.mod = TRUE)) #the ricker.mod is a modification that is used within the equation
summary(leslie) 
#remember that N0 = initial population size and q = catchability coefficient or the fraction of the population that is removed by one unit of fishing effort

#find the confidence intervals of this data
confint(leslie)

#plot it
plot(leslie)
```
### The initial population size of the Pristipomoides zonatus is 1066. 

## Now for the DeLury Method!
```{r}
#we will use the same set of data and species for this method in order to compare/contrast them
delury <- with(snapper_pzonatus, depletion(pzonatus, effort, method="Delury", Ricker.mod = TRUE))
summary(delury)

confint(delury)

plot(delury)
```
### The DeLury Method finds that the inital population size of Pristipomoides zonatus is 1077, so 11 more than the Leslie Method as well as a slight decrease in q!

## Now for the K-Pass Method
```{r}
#again, we are going to use the snapper Pristipomoides zonatus data, but this time we only need catch data
#for this method, we will still get N0 but will also get p which is probability of capture

k_pass <- with(snapper_pzonatus, removal(pzonatus)) 
summary(k_pass)

confint(k_pass)
```
### The K-Pass Method determined that N0 is 985, but does have an upper bound of 1077. p=0.1

Portfolio Assignment: Grab a new set of data and use the Leslie, DeLury, and K-Pass methods to calculate the initial population size! Answer the following questions:

1. Compare and contrast the initial population sizes between the three methods
2. Why are the Leslie and DeLury methods not valid at estimating N0? 
3. What values can you look at to determine the reliability of the data? (Hint: Think back to what we learned about statistics in 206)
